Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2010-12-28 rebuttal.tex

We find other suggestions on presentation useful and have made changes accordingly.

\paragraph{About Positioning and Novelty:}
As we have tried to make it clear in the very beginning, the proposed method does not aim to solve face 
recognition for all scenarios! It is not designed for the case where the training is uncontrolled and limited, hence it
may not have any advantages over other face recognition techniques/systems specifically designed for such scenarios. 
However, there are many practical scenarios where the proposed setting is very relevant: access control
or recognizing refugees, where cooperative training images can (and should) be obtained in advance but no cooperative test
images are needed or available. No need of cooperative test images is arguably the main (or only) advantage that 
face recognition has over other biometric measures for security, such as fingerprint, iris, and DNA. 
If the subjects can cooperate for both training and testing, as suggested by the reviewers, then face 
recognition may no longer be the best choice, in terms of cost and accuracy. The proposed system is an attempt
to build a system that is not only accurate and robust enough to compete with other biometric measures but also 
discreet and non-intrusive when in use. To the best of our knowledge (probably the reviewer may agree), even 
for this admittedly restricted scenario, there is no practical face-recognition system whose performance can meet
the standards of most access control and security purposes. Existing systems have found success  mostly in applications
that are less-demanding in performance such as online image search or photo album organization etc.

Thanks to the reviewer's comments, we have modified the introduction to make this point more clear, especially clarifying the specific scenario where our method can claim novelty.

@@

\paragraph{About Summary Recommendation:} While we agree with the reviewer that careful evaluation of our
system is needed by comparing it with other face recognition techniques/systems. We must point out that most face recognition systems are not designed for the scenario considered by this paper, hence a fair and meaningful comparison can be difficult. Nevertheless, to better evaluate both the strength and limitation of our method, we have added experimental comparison of our system with LBP, one of the leading techniques used for face verification for most commercial face recognition systems (in addition to other techniques already compared in the paper). We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
datasets.  The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is not controlled nor sufficient.

\paragraph{Detailed Concerns or Comments by Reviewer 1:}
\begin{enumerate}
\item As we have mentioned earlier, we completely agree with the reviewer that if the users are completely cooperative at
both training and testing phases, face recognition may not necessarily be the best choice of biometric measure -- fingerprint
and iris are arguably better choices. However, our method does NOT require the users to be completely cooperative at the test stage. The purpose of having a good training set and a robust recognition system is precisely to alleviate the users 
from being so cooperative at the test stage. This will allow the system to be less intrusive and much more user-friendly 
as it can handle typically daily change of faces such as occlusion by glasses, hats, or beard and variation in lighting and shadow. Ideally, face recognition will take place automatically and unknowingly as a user walks through a door or works in front a desktop.

@@

\begin{enumerate}
\item (i) ``How often an infinite loop is encountered, and how it is handled."

We place a hard limit on the maximal number of alignment iterations, so the
algorithm is guaranteed to terminate in finite time.  In practice, the
alignment algorithm tends to converge within 4-10 iterations for each
resolution of the multi-scale implementation.  We also place hard bounds on how
large the alignment can be, and terminate alignment early if these bounds are
reached.  In the cases where the alignment ``slips off'' the face, this is how
alignment terminates.

\item (ii) ``How often the starting point happens to be located in a wrong basin of
attraction, i.e. a wrong face. (If there is any way to know this.)''

Since we align the test image to the training users individually, we avoid the
potential problem of the test image converging to a local minimum corresponding
to a wrong face.

@@

\item (iii) ``How often the solution gets drifted away into a wrong basin of
attraction. (If there is any way to know this.)''

In Section II E, we have conducted careful experiments to measure the region of attraction
for the proposed alignment method. We have studied the region of attraction of the proposed
method for both 2D and 3D pose variation, as well as how it can be improved with a proper
multi-scale implementation.

@@

We have attempted to make the overall structure of our system clear earlier in the
paper, by clarifying early on that SRC forms the recognition stage of our system
in the introduction, as well as by making it clear in the caption of the motivating example
is not the final pipeline.

\item ``Using and requiring several input images of a person asks for a
comment on the advantage or difference to a system that uses
a 3D representation of a face. Please comment on this!"

The primary advantage of systems that internally use a 3D model of human faces
is that they have the potential to work over a wider range of pose variation in
the test image.  In general, systems using 2D models are simpler and easier to
engineer.  As far as data acquisition is concerned, 2D cameras still have a far
higher resolution and lower noise levels than 3D cameras.  Based on the principles
of photometric stereos, our set of 2D images may very well encode more useful shape 
information about a user's face than a single low-resolution depth image, especially if 
recognition is to be performed on 2D input images.

In addition, the sparse representation techniques are not limited to frontal 2D images.
They can be easily applied to 2D images from multiple poses or to establish alignment
between a 2D image and a 3D face model (if available). These are all good directions
for future research.

We have modified Section I.A and the Conclusion to address these issues.
